---
title: "Lab 9 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Seating
1. If you already have a group, it will be helpful to start sitting next to each other.
2. If you don't have a group, hang tight we are working on it.
3. Please setup your computer as normal for today.

## Warm-up
1. In the data folder there is an epidemiology data set on an outbreak of malaria.
```{r}
library(tidyverse)
```
```{r}
malaria<- readr::read_csv("data/malaria.csv") %>% clean_names()
malaria
```

2. `rdt` refers to rapid diagnostic test and they are identified here by age group.


3. Make the data tidy and store them as a new object.
```{r}
malaria_long <- 
  malaria %>% 
  pivot_longer(
    cols = starts_with("malaria"),
    names_to = "age_range",
    values_to = "cases"
  ) %>% 
  select(newid, data_date, submitted_date, province, district, location_name, age_range, cases)
malaria_long
```

4. Which district had the highest *total* number of cases on July 30, 2020?

```{r}
malaria_long %>% 
  filter(data_date == "2020-07-30") %>%
  group_by(district) %>% 

```

